---
title: "Little Lost - Pahsimeroi - Winter"
output:
  html_document:
    df_print: paged
---
## Winter
Notes: will need to make some weight adjustments based on LMF segment size

### Setup
Set local library path and load required packages
```{r setup, eval =FALSE}
library(aim.analysis)
library(sp)
library(dplyr)
library(binom)
library(stringr)
```

### Defaults and filepaths

```{r}
# Projection to use. Alber's Equal Area is useful for calculating areas
projection <- sp::CRS("+proj=aea")

# Date of analysis
date <- Sys.Date()

# Confidence level in percent
confidence <- 80

output_path <- "C:\\Users\\alaurencetraynor\\Documents\\2019\\Analysis\\Pahsimeroi\\LittleLostPaVa\\2020\\outputs"

# Filename for ratings from analysis requestor
ratings_filename <- "HAF_ID_2020_LittleLost_Pahsimeroi_SiteScale_RESULTS_02-05-2020.xlsx"

# Core filename for all outputs
output_filename <- "pava_winter"

# Reporting Unit name - corresponds to column header in ratings spreadsheet
sua_name <- "Suitability: \r\nWinter-S6"

# The geodatabase and layer names for the spatial data sources
# are embedded in the code below

# Habitat type (corresponds to sheet name in ratings Excel workbook)
# this worksheet has all three SUA ratings so need to filter out unnecessary columns
seasonal_habitat_type = "Plot Summary"

# The variable in AIM sample designs that contains the unique identifier for each point
aim_idvar <- "PLOT_NM"
# The variable in AIM sample designs that contains the fate for each point
aim_fatevar = "FINAL_DESIG"
# The variable in TerrADat that contains the unique identifier for each point matching AIM designs'
tdat_idvar <- "PlotID"
# The variable in LMF points that contains the unique identifier for each point
lmf_idvar <- "PLOTKEY"

# The fates in sample designs which correspond to sampled/observed
observed_fates = c("TS", "Target Sampled", "Sampled")

# The fates in sample designs which correspond to unneeded
# (e.g. unused oversample or points designated for future sampling)
invalid_fates = c("NN", NA, "Non Needed", "", "Not Sampled")

# The LMF segments polygons
# The current workflow is to join LMF segments to points in ArcMap so segment polys are only needed to calculate adjusted weights for segments which have been cut in half
segments <- NULL

# The variable containing the LMF segment identity/membership
# This is found in the points and the segment polygons
segment_var <- "segment_id"

# The variable that contains the areas in HECTARES in wgtcats
# This is calculated in ArcMap
wgtcat_area_var = "area_hectares"

# The unique ID for poststratification polygons
wgtcats_var <- "wgtcat"
```

### The filepath to the folder where we have all the spatial data
```{r}
path_spatial <- "C:\\Users\\alaurencetraynor\\Documents\\2019\\Analysis\\Pahsimeroi\\LittleLostPaVa\\2020\\inputs\\spatial"
```
### The filepath to the folder where we have the Excel stuff
```{r}
path_tabular <- "C:\\Users\\alaurencetraynor\\Documents\\2019\\Analysis\\Pahsimeroi\\LittleLostPaVa\\2020\\inputs\\tabular"
```

### Read in
#### Reporting unit(s)
Strata and points are already clipped to reporting units in Arc

#### Strata polygons
```{r}
wgtcat_df <-sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                        layer = "LittleLost_Pahsimeroi_Winter_Strata",
                        stringsAsFactors = FALSE)
sf::st_geometry(wgtcat_df) <- NULL
wgtcats <- as.data.frame(wgtcat_df)[,c(1:3)]
```

#### Plot Ratings
```{r}
ratings <- readxl::read_xlsx(paste(path_tabular, ratings_filename, sep = "/"),
                             sheet = seasonal_habitat_type)

ratings <- subset(ratings[,c("Project","Plot ID", "Visit Date",sua_name)], !is.na(ratings[[sua_name]]) & ratings[[sua_name]] != "" & ratings[[sua_name]] != "0") # we are only interested in those plots rated in our sua of interest
```

#### Sampled Points 
Make sure points and polygons do not include Z/M dimensions...
```{r}
sampled_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "LMF_TerrADat_segments_strata_join_WinterRated",
                           stringsAsFactors = FALSE)
## Standardize projection
sampled_points <- sf::st_transform(sampled_points,
                                crs = projection)
## Make spatial 
sampled_points <- methods::as(sampled_points, "Spatial")

## sampled plots joined to LMF segments in ArcGIS

## standardise wgtcat variable name
sampled_points@data[["wgtcat"]] <- sampled_points@data[[wgtcats_var]]
```

#### Design points
```{r}
### Challis
design1_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "design_points_challis_winter_join",
                           stringsAsFactors = FALSE)
## Standardize projection
design1_points <- sf::st_transform(design1_points,
                                crs = projection)
## Make spatial 
design1_points <- methods::as(design1_points, "Spatial")

# Repeat for each sample design 
### Uppersnake
design2_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "design_points_uppersnake_winter_join",
                           stringsAsFactors = FALSE)
## Standardize projection
design2_points <- sf::st_transform(design2_points,
                                crs = projection)
## Make spatial 
design2_points <- methods::as(design2_points, "Spatial")
```

###### Sort out points ######
#### Prepare TerrADat points
```{r}
### Make naming consistent across all data
# Move unique IDs to unique_id in TerrADat
sampled_points@data[["unique_id"]] <- sampled_points@data[[tdat_idvar]]
# Set fate variable in TerrADat
sampled_points@data[["fate"]] <- observed_fates[1]

# Move LMF segment membership to segment
sampled_points@data[["segment"]] <- sampled_points@data[[segment_var]]

# repeat for all points
design1_points@data[["unique_id"]] <- design1_points@data[[aim_idvar]]
design1_points@data[["fate"]] <- design1_points@data[[aim_fatevar]]
design1_points@data[["wgtcat"]] <- design1_points@data[[wgtcats_var]]
design1_points@data[["segment"]] <- design1_points@data[[segment_var]]

design2_points@data[["unique_id"]] <- design2_points@data[[aim_idvar]]
design2_points@data[["fate"]] <- design2_points@data[[aim_fatevar]]
design2_points@data[["wgtcat"]] <- design2_points@data[[wgtcats_var]]
design2_points@data[["segment"]] <- design2_points@data[[segment_var]]
```


#### Create combined points shapefile
### Combine all sample design points
```{r}
### Mash the design files together, keeping only necessary variables
design_points <- rbind(design1_points[, c("unique_id", "fate", "wgtcat", "segment")],
                       design2_points[, c("unique_id", "fate", "wgtcat", "segment")]) 

sum(sampled_points$unique_id %in% design_points$unique_id) # number of sampled AIM plots also present in design
sum(design_points$unique_id[design_points$fate %in% observed_fates] %in% sampled_points$unique_id) # number of observed desing points in sampled points
```
### QC
```{r}
# filter out LMF
# separate out lmf point for code below
lmf_points <- sampled_points
lmf_points <- subset(lmf_points, is.na(lmf_points[[tdat_idvar]]))
lmf_points[["unique_id"]] <- lmf_points[[lmf_idvar]]
sampled_points <- subset(sampled_points, !is.na(sampled_points@data$unique_id))
# Add in coordinates (needed for the analysis step)
lmf_points <-  aim.analysis::add_coords(lmf_points,
                                        xynames = c("x", "y"))

# Checking against the ratings confirms if all the rated plots are represented (important)
# and if all the sampled_plots are rated (not important, but informative)
if (!all(ratings[["Plot_ID"]][ratings[["AIM"]] == "AIM"] %in% sampled_points$unique_id)) {
  warning("NOT ALL RATED AIM PLOTS ARE PRESENT IN sampled_points!")
}
all(sampled_points$unique_id %in% ratings[["Plot ID"]]) # this is just looking at AIM points at this point

# filter sampled plots by those that were rated
# This should have been done in Arc GIS first
# sampled_points <- sampled_points[sampled_points$unique_id %in% ratings[["Plot ID"]], ] 

```
# Now combine:
# All design_points which were not observed. This removes any points that overlap with sampled_points but also
# any that were observed but somehow not in TerrADat and deemed unsuitable for rating
# which therefore means they also should not be included in weighting
# All sampled points (already restricted to the ones which were rated)

```{r}
aim_points <- rbind(design_points[!(design_points$fate %in% observed_fates), c("unique_id", "fate", "wgtcat", "segment")], 
                    sampled_points[, c("unique_id", "fate", "wgtcat", "segment")])

# Take out invalid fates, e.g. things that don't get included in weighting
aim_points <- aim_points[!(aim_points$fate %in% invalid_fates), ]

# Add in coordinates (needed for the analysis step)
aim_points <-  aim.analysis::add_coords(aim_points,
                                        xynames = c("x", "y"))

aim_points$wgtcat <- as.character(aim_points$wgtcat) 
```

###### Analyze ######
############################################################
#### Calculate weights ####
### Note that this is the contents of aim.analysis::weight_aimlmf() verbatim
### As a function, it was producing errors that it does not as bare code
############################################################

This functions is giving errors:
aim.analysis::weight_aimlmf(aim_points = aim_points,
                            lmf_points = lmf_points,
                            aim_idvar <- "unique_id",
                            lmf_idvar <- "unique_id",
                            aim_fatevar <- "fate",
                            wgtcat_var <- wgtcats_var,
                            segments <- NULL,
                            segment_var <- "segment",
                            verbose <- FALSE )
Using raw code instead:                     
```{r}
# These are being redefined here as though they were arguments to the function
aim_idvar <- "unique_id"
lmf_idvar <- "unique_id"
aim_fatevar <- "fate"
wgtcat_var <- "wgtcat"
segments <- NULL
segment_var <- "segment"
verbose <- FALSE

if (length(wgtcat_var) != 1 | class(wgtcat_var) != "character") {
  stop("wgtcat_var must be a single character string")
}
if (!is.null(wgtcat_area_var)) {
  if (length(wgtcat_area_var) != 1 | class(wgtcat_area_var) != "character") {
    stop("wgtcat_area_var must be a single character string")
  }
}
if (length(segment_var) != 1 | class(segment_var) != "character") {
  stop("segment_var must be a single character string")
}

if (!is.null(wgtcats)) {
  if (!(class(wgtcats) %in% c("SpatialPolygonsDataFrame", "data.frame"))) {
    stop("wgtcats must be a spatial polygons data frame or data frame")
  }
  if (nrow(wgtcats) < 1) {
    stop("wgtcats contains no observations/data")
  }
  if (!(wgtcat_var %in% names(wgtcats))) {
    stop(paste("The variable", wgtcat_var, "does not appear in wgtcats@data"))
  }
}

if (!is.null(segments)) {
  if (!(class(segments) %in% "SpatialPolygonsDataFrame")) {
    stop("segments must be a spatial polygons data frame")
  }
  if (nrow(segments) < 1) {
    stop("segments contains no observations/data")
  }
  if (!(segment_var %in% names(segments))) {
    stop(paste("The variable", segment_var, "does not appear in segments@data"))
  }
}

if (is.null(aim_fatevar)) {
  warning("No fate variable specified for AIM points. Assuming all were observed/sampled.")
  aim_fatevar <- "fate"
  aim_points[["fate"]] <- "observed"
  lmf_points[["fate"]] <- "observed"
  observed_fates <- "observed"
} else {
  if (length(aim_fatevar) > 1 | class(aim_fatevar) != "character") {
    stop("The aim fate variable must be a single character string")
  }
  if (!aim_fatevar %in% names(aim_points)) {
    stop(paste("The variable", aim_fatevar, "does not appear in aim_points@data"))
  } else {
    aim_points[["fate"]] <- aim_points[[aim_fatevar]]
  }
  if (is.null(observed_fates)) {
    warning("No observed fates provided. Assuming all AIM points were observed/sampled unless specified otherwise with invalid_fates")
    observed_fates <- unique(aim_points[["fate"]])
    observed_fates <- observed_fates[!(observed_fates %in% invalid_fates)]
  }
}

lmf_points[["fate"]] <- observed_fates[1]

if (!is.null(observed_fates)) {
  if (!any(aim_points[["fate"]] %in% observed_fates)) {
    warning("No AIM points have a fate specified as observed.")
  }
}

# Harmonize projections
if (is.null(projection)) {
  if (!is.null(wgtcats)) {
    projection <- wgtcats@proj4string
  } else if (!is.null(segments)) {
    projection <- segments@proj4string
  }
}

if (!is.null(projection)) {
  if (class(aim_points) %in% c("SpatialPointsDataFrame")) {
    if (!identical(projection, aim_points@proj4string)) {
      aim_points <- sp::spTransform(aim_points,
                                    CRSobj = projection)
    }
  }
  if (class(lmf_points) %in% c("SpatialPointsDataFrame")) {
    if (!identical(projection, lmf_points@proj4string)) {
      lmf_points <- sp::spTransform(lmf_points,
                                    CRSobj = projection)
    }
  }
  if (!is.null(wgtcats)) {
    if (class(lmf_points) %in% c("SpatialPolygonsDataFrame")) {
      if (!identical(projection, wgtcats)) {
        wgtcats <- sp::spTransform(wgtcats,
                                   CRSobj = projection)
      }
    }
  }
  if (!is.null(segments)) {
    if (!identical(projection, segments@proj4string)) {
      segments <- sp::spTransform(segments,
                                  CRSobj = projection)
    }
  }
}



# Assign the weight categories
if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
  if (!(wgtcat_var %in% names(wgtcats))) {
    stop("The variable ", wgtcat_var, " does not appear in wgtcats")
  }
  aim_points[["wgtcat"]] <- sp::over(aim_points, wgtcats)[[wgtcat_var]]
  lmf_points[["wgtcat"]] <- sp::over(lmf_points, wgtcats)[[wgtcat_var]]
} else {
  if (!(wgtcat_var %in% names(aim_points))) {
    stop("The variable ", wgtcat_var, " does not appear in aim_points")
  }
  if (!(wgtcat_var %in% names(lmf_points))) {
    stop("The variable ", wgtcat_var, " does not appear in lmf_points")
  }
  aim_points[["wgtcat"]] <- aim_points[[wgtcat_var]]
  lmf_points[["wgtcat"]] <- lmf_points[[wgtcat_var]]
}


# Assign the LMF segment codes
if (is.null(segments)) {
  if (!(segment_var %in% names(aim_points))) {
    stop("The variable ", segment_var, " does not appear in aim_points")
  }
  if (!(segment_var %in% names(lmf_points))) {
    stop("The variable ", segment_var, " does not appear in lmf_points")
  }
  aim_points[["segment"]] <- aim_points[[segment_var]]
  lmf_points[["segment"]] <- lmf_points[[segment_var]]
} else {
  if (!(segment_var %in% names(segments))) {
    stop("The variable ", segment_var, " does not appear in segments")
  }
  aim_points[["segment"]] <- sp::over(aim_points, segments)[[segment_var]]
  lmf_points[["segment"]] <- sp::over(lmf_points, segments)[[segment_var]]
}


# Just harmonize the idvar names for now
aim_points[["unique_id"]] <- aim_points[[aim_idvar]]
lmf_points[["unique_id"]] <- lmf_points[[lmf_idvar]]

# Add reporting units
aim_points[["reporting_unit"]] <- seasonal_habitat_type
lmf_points[["reporting_unit"]] <- seasonal_habitat_type

# If somehow LMF points aren't in a segment, that's a major problem
if (any(is.na(lmf_points[["segment"]]))) {
  stop(paste("The following LMF points did not spatially intersect any segment polygons:",
             paste(lmf_points[is.na(lmf_points[["segment"]]), "unique_id"], collapse = ", ")))
}

# TODO: Stick a check in here that the segment ID from the polygons
# matches the one derived from the LMF plot ID
# Probably just warn if not?


# Get data frames
if (class(aim_points) %in% "SpatialPointsDataFrame") {
  aim_df <- aim_points@data
} else {
  aim_df <- aim_points
}
if (class(lmf_points) %in% "SpatialPointsDataFrame") {
  lmf_df <- lmf_points@data
} else {
  lmf_df <- lmf_points
}

# NOTE THAT THIS FILTERS OUT ANYTHING FLAGGED AS NOT NEEDED IN THE FATE
# So that'd be unused oversamples or points from the FUTURE that no one would've sampled anyway
aim_df <- aim_df[!(aim_df[["fate"]] %in% invalid_fates), c("unique_id", "fate", "wgtcat", "segment")]
aim_df[["aim"]] <- TRUE
aim_df[["lmf"]] <- FALSE

# We only have target sampled LMF points available to us, so we don't need to filter them
lmf_df <- lmf_df[, c("unique_id", "fate", "wgtcat", "segment")]
lmf_df[["aim"]] <- FALSE
lmf_df[["lmf"]] <- TRUE

# Combine them
combined_df <- unique(rbind(aim_df, lmf_df))

# There shouldn't be any that don't belong to a wgtcat anymore
combined_df <- combined_df[!is.na(combined_df[["wgtcat"]]), ]

# Add an observed variable for easy reference later
combined_df[["observed"]] <- combined_df[["fate"]] %in% observed_fates

# To make the lookup table, drop any points that fell outside LMF segments
combined_segmentsonly_df <- combined_df[!is.na(combined_df[["segment"]]), ]

# Create a segment relative weight lookup table
segment_relwgt_lut <- do.call(rbind,
                              lapply(X = split(combined_segmentsonly_df, combined_segmentsonly_df[["segment"]]),
                                     FUN = function(X){
                                       # These are the count of AIM points with any valid fate
                                       aim_count <- sum(X[["aim"]])
                                       
                                       # We also need the count of LMF points with any valid fate, but that's complicated
                                       # We only have the sampled LMF points so we can count those
                                       lmf_sampled_count <- sum(X[["lmf"]])
                                       # To get the number of evaluated but not sampled points:
                                       # The LMF plot keys end in a digit that represents the intended sampling order within a segment
                                       # 1 and 2 are considered base points and were intended to be sampled
                                       # If a sampled LMF plot's plot key ends in 3, that means that one or both of the base points
                                       # were evaluated and rejected rather than sampled, which brings the evaluated LMF plot count
                                       # to three for the segment.
                                       # This just asks if the third point was used
                                       lmf_oversample_used <- any(grepl(X[["unique_id"]][X[["lmf"]]],
                                                                        pattern = "\\D3$"))
                                       
                                       # Likewise, if only one LMF plot was sampled in a segment, that means the other two were
                                       # evalurated and rejected rather than sampled, also bringing the total to three.
                                       # So if there was only one sampled or if the oversample was used, there were three evaluated
                                       if (sum(X[["lmf"]]) == 1 | lmf_oversample_used) {
                                         lmf_count <- 3
                                       } else {
                                         # This will fire only if there sampled count was 2, but better to be safe here
                                         lmf_count <- lmf_sampled_count
                                       }
                                       
                                       
                                       # The relative weight for points falling within a segment is calculated as
                                       # 1 / (number of points)
                                       relative_weight <- 1 / sum(aim_count, lmf_count)
                                       
                                       output <- data.frame("segment" = X[["segment"]][1],
                                                            "relwgt" = relative_weight,
                                                            stringsAsFactors = FALSE)
                                       return(output)
                                     }))

# Add the relative weights to the combined AIM and LMF points
combined_df <- merge(x = combined_df,
                     y = segment_relwgt_lut,
                     by = "segment",
                     all.x = TRUE)

# Anywhere there's an NA associated with an AIM point, that's just one that fell outside the segments
combined_df[is.na(combined_df[["relwgt"]]) & combined_df[["aim"]], "relwgt"] <- 1

# Correct relwgt for LMF segment cut in half
combined_df[combined_df[["unique_id"]]=="20141637102304B2","relwgt"] <- combined_df[combined_df[["unique_id"]]=="20141637102304B2","relwgt"] *0.5

if (is.null(wgtcat_area_var)) {
  if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
    wgtcat_df <- aim.analysis::add.area(wgtcats)@data
  } else {
    stop("No name for a variable in wgtcats containing the area in hectares was provided and wgtcats is not a spatial polygons data frame so area can't be calculated")
  }
} else {
  if (!(wgtcat_area_var %in% names(wgtcats))) {
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- aim.analysis::add.area(wgtcats)@data
    } else {
      stop("The variable ", wgtcat_area_var, " does not appear in wgtcats")
    }
  } else {
    warning("Trusting that the variable ", wgtcat_area_var, " in wgtcats contains the areas in hectares")
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- wgtcats@data
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    } else {
      wgtcat_df <- wgtcats
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    }
  }
}


wgtcat_df[["wgtcat"]] <- wgtcat_df[[wgtcat_var]]

# Making sure that these aren't spread out over multiple observations
if (verbose) {
  message("Summarizing wgtcat_df by wgtcat to calculate areas in case the wgtcats are split into multiple observations")
}
wgtcat_df <- dplyr::summarize(dplyr::group_by(wgtcat_df,
                                              wgtcat),
                              "hectares" = sum(AREA.HA))

wgtcat_areas <- setNames(wgtcat_df[["hectares"]], wgtcat_df[["wgtcat"]])

# had a issue here reading wgtcats as integers
wgtcat_df$wgtcat <-  as.character(wgtcat_df$wgtcat)
  
weight_info <- lapply(X = wgtcat_df[["wgtcat"]],
                      wgtcat_areas = wgtcat_areas,
                      points = combined_df,
                      wgtcat_var = wgtcat_var,
                      FUN = function(X, wgtcat_areas, points, wgtcat_var){
                        # Area of this polygon
                        area <- wgtcat_areas[X]
                        
                        # All of the points (AIM and LMF) falling in this polygon
                        points <- points[points[["wgtcat"]] == X, ]
                        
                        # If there are in fact points, do some weight calculations!
                        if (nrow(points) > 0 & any(points[["observed"]])) {
                          # The number of observed AIM points with a relative weight of 1
                          # So, not sharing an LMF segment with any LMF points
                          aim_standalone <- sum(combined_df[["aim"]] & combined_df[["relwgt"]] == 1)
                          # The number of unique segments selected for LMF points
                          lmf_segment_count <- length(unique(points[["segment"]]))
                          
                          # The approximate number of 160 acre segments in this polygon
                          # Obviously, not all of the segments were selected in the first stage
                          # of the LMF design, but this is how many were available in this polygon
                          approximate_segment_count <- area / (160 / 2.47)
                          
                          # This is the sum of the relative weights of the OBSERVED points
                          # This does not include the inaccessible, rejected, or unknown points
                          # It does include both AIM and LMF, however
                          sum_observed_relwgts <- sum(points[points[["observed"]], "relwgt"])
                          # This is the sum of the relative weights of all the AIM points, regardless of fate
                          sum_relwgts <- sum(points[["relwgt"]])
                          
                          # The units are segments per point
                          # The segments are 120 acre quads (quarter sections?) that the first stage of LMF picked from
                          # Steve Garman called this "ConWgt" which I've expanded to conditional_weight
                          # but that's just a guess at what "con" was short for
                          conditional_weight <- approximate_segment_count / (aim_standalone + lmf_segment_count)
                          conditional_area <- conditional_weight * sum_observed_relwgts
                          
                          # What's the observed proportion of the area?
                          observed_proportion <- sum_observed_relwgts / sum_relwgts
                          # And how many acres is that then?
                          # We can derive the "unknown" or "unsampled" area as the difference
                          # between the polygon area and the observed area
                          observed_area <- area * observed_proportion
                          
                          # Then this adjustment value is calculated
                          weight_adjustment <- observed_area / conditional_area
                          
                          
                          
                          # Put everything about the wgtcat in general in one output data frame
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = approximate_segment_count,
                                                      sum_observed_relwgts = sum_observed_relwgts,
                                                      sum_relwgts = sum_relwgts,
                                                      observed_proportion = observed_proportion,
                                                      observed_area = observed_area,
                                                      unobserved_area = area - observed_area,
                                                      conditional_weight = conditional_weight,
                                                      conditional_area = conditional_area,
                                                      weight_adjustment = weight_adjustment,
                                                      point_count = sum(points[["observed"]]),
                                                      observed_point_count = nrow(points),
                                                      stringsAsFactors = FALSE)
                          
                          # But much more importantly add the calculated weights to the points
                          output_points <- points
                          
                          # This handles situations where there were points, but none of them were observed
                          # In that case, weight_adjustment will be 0 / 0 = NaN
                          # That's fine because we can identify that this polygon is still in the inference area
                          # but that its entire area is "unknown" because no data were in it
                          if (is.nan(weight_adjustment)) {
                            output_points[["wgt"]] <- NA
                          } else {
                            # The point weight is calculated here.
                            # This is Garman's formula and I don't have the documentation justifying it on hand
                            output_points[["wgt"]] <- weight_adjustment * conditional_weight * points[["relwgt"]]
                            message("Checking weight sum for ", X)
                            # I'm rounding here because at unrealistically high levels of precision it gets weird and can give false positives
                            if (round(sum(output_points[["wgt"]]), digits = 3) != round(area, digits = 3)) {
                              warning("The sum of the point weights (", sum(output_points[["wgt"]]), ") does not equal the polygon area (", area, ") for ", X)
                            }
                          }
                          
                        } else {
                          # Basically just empty data frames
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = area / (160 / 2.47),
                                                      sum_observed_relwgts = NA,
                                                      sum_relwgts = NA,
                                                      observed_proportion = 0,
                                                      observed_area = 0,
                                                      unobserved_area = area,
                                                      conditional_weight = NA,
                                                      conditional_area = NA,
                                                      weight_adjustment = NA,
                                                      point_count = 0,
                                                      observed_point_count = 0,
                                                      stringsAsFactors = FALSE)
                          output_points <- NULL
                        }
                        
                        return(list(points = output_points,
                                    wgtcat = output_wgtcat))
                      })

point_weights <- do.call(rbind,
                         lapply(X = weight_info,
                                FUN = function(X){
                                  X[["points"]]
                                }))
wgtcat_summary <- do.call(rbind,
                          lapply(X = weight_info,
                                 FUN = function(X){
                                   X[["wgtcat"]]
                                 }))

```

###### Analysis ######
```{r}
all_points <- rbind(aim_points[, c("unique_id", "x", "y")],
                    lmf_points[, c("unique_id", "x", "y")])

# Add the ratings to the point info and convert to a data frame
all_points <- merge(x = all_points@data,
                    y = ratings,
                    by.x = "unique_id",
                    by.y = "Plot ID",
                    all.x = FALSE)

# Add in the indicator, which for these will always be "Habitat suitability"
all_points[["indicator"]] <- "Habitat suitability"
all_points[[paste(sua_name)]] <- as.character(all_points[[paste(sua_name)]]) # analyze function giving "lossy cast error" - need to convert factors to characters

# Add in the season (this is defined up above) e.g. "S-4 Upland Summer Late Brood-rearing Habitat"
all_points[["reporting_unit"]] <- paste(sua_name)
point_weights[["reporting_unit"]] <- paste(sua_name)

# And analyze
analysis <- aim.analysis::analyze(benchmarked_points = all_points,
                                  point_weights = point_weights,
                                  id_var = "unique_id",
                                  indicator_var = "indicator",
                                  value_var = sua_name,
                                  x_var = "x",
                                  y_var = "y",
                                  reporting_var = "reporting_unit",
                                  weight_var = "wgt",
                                  conf = confidence)
```

### Confidence Interval Function

```{r}
goodman_cis <- function(counts,
                        alpha = 0.2,
                        chisq = "best",
                        verbose = FALSE){
  
  if (!(chisq %in% c("A", "B", "best"))) {
    stop("The only valid values for chisq are 'A', 'B', and 'best'.")
  }
  
  # Goodman describes the upper and lower bounds with the equations:
  # Lower estimated pi_i = {A + 2n_i - {A[A + 4n_i(N - n_i) / N]}^0.5} / [2(N + A)]
  # Upper estimated pi_i = {A + 2n_i + {A[A + 4n_i(N - n_i) / N]}^0.5} / [2(N + A)]
  
  # n_i is the "observed cell frequencies in population of size N" (aka count of observations) from a category
  # so that's the incoming argument counts. We'll rename for consistency with the original math (and statistics as a discipline)
  n <- counts
  
  # N is the population those counts make up, or, in lay terms, the total observation count
  N <- sum(counts)
  
  # k is the number of categories the population has been sorted into
  # Useful for degrees of freedom
  k <- length(counts)
  
  # "A is the upper alpha * 100-th percentage point of the chi-square distribution with k - 1 degrees of freedom"
  # and B is an alternative which uses alpha / k and one degree of freedom
  # Goodman states that B should be less than A for situations
  # where k > 2 AND alpha is 0.1, 0.05, or 0.01.
  chisq_quantiles <- c("A" = stats::qchisq(p = 1 - alpha,
                                           df = k - 1),
                       "B" = stats::qchisq(p = 1 - (alpha / k),
                                           df = 1))
  
  
  # According to Goodman, A and B are both valid options for the chi-square quantile
  # So the user can specify which they want or just ask for the one that minimizes the confidence intervals
  chisq_quantile <- switch(chisq,
                           "A" = {chisq_quantiles["A"]},
                           "B" = {chisq_quantiles["B"]},
                           "best" = {
                             pick <- which.min(chisq_quantiles)
                             if (verbose){
                               switch(names(chisq_quantiles)[pick],
                                      "A" = message("The chi-square quantile calculation that will provide the tighter confidence intervals is A, the upper alpha X 100-th percentage point of the chi-square distribution with k - 1 degrees of freedom"),
                                      "B" = message("The chi-square quantile calculation that will provide the tighter confidence intervals is B, the upper alpha / k X 100-th percentage point of the chi-square distribution with 1 degree of freedom"))
                             }
                             chisq_quantiles[pick]
                           })
  
  # Calculate the bounds!
  # Note that these ARE symmetrical, just not around the proportions.
  # They're symmetrical around chisq_quantile + 2 * n / (2 * (N + chisq_quantile))
  # The variable A has been replaced with chisq_quantile because it may be A or B, depending
  # Since the only multi-value vector involved here is n, these will be vectors of length k,
  # having one value for each of the values in n and in the same order as n
  lower_bounds <- (chisq_quantile + 2 * n - sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))) / (2 * (N + chisq_quantile))
  upper_bounds <- (chisq_quantile + 2 * n + sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))) / (2 * (N + chisq_quantile))
  
  # A proportion can never be greater than 1 or less than 0 (duh)
  # That's definitely a thing that can happen if the magnitude of sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))
  # is large enough
  lower_bounds[lower_bounds < 0] <- 0
  upper_bounds[upper_bounds > 1] <- 1
  
  # Build the output
  output <- data.frame(count = n,
                       proportion = n / N,
                       lower_bound = lower_bounds,
                       upper_bound = upper_bounds)
  
  # What are the categories called? If anything, that is
  k_names <- names(n)
  
  if (!is.null(k_names)) {
    output[["category"]] <- k_names
    output <- output[, c("category", "count", "proportion", "lower_bound", "upper_bound")]
  }
  
  return(output)
}

```

### Formatting Outputs
```{r}
analysis <- analysis[, c("Subpopulation",
                         "Indicator",
                         "Category",
                         "NResp",
                         "Estimate.P",
                         "Estimate.U")]

adjusted_counts <- (sum(analysis[["NResp"]][analysis[["Category"]] != "Total"]) * analysis[["Estimate.P"]][analysis[["Category"]] != "Total"] / 100)

goodman_cis <- goodman_cis(counts = adjusted_counts)

goodman_cis[["Category"]] <- analysis[["Category"]][analysis[["Category"]] != "Total"]

analysis <- merge(x = analysis,
                  y = goodman_cis[, c("Category", "lower_bound", "upper_bound")])

analysis[["lower_bound"]] <- analysis[["lower_bound"]] * 100
analysis[["upper_bound"]] <- analysis[["upper_bound"]] * 100

# Calculate the hectares using the Goodman binomial confidence intervals from percentage
total_hectares <- analysis[["Estimate.U"]]/(analysis[["Estimate.P"]]/100)
  
analysis[[paste0("LCB", confidence, "Pct.U.Goodman")]] <- analysis[["lower_bound"]] / 100 * total_hectares
analysis[[paste0("UCB", confidence, "Pct.U.Goodman")]] <- analysis[["upper_bound"]] / 100 * total_hectares

names(analysis) <- c("Rating",
                     "Habitat Type",
                     "Indicator",
                     "Number of plots",
                     "Estimated percent of sampled area",
                     "Estimated hectares",
                     paste0(c("Lower confidence bound of percent area (", "Upper confidence bound of percent area ("), confidence, "%, Goodman multinomial) "),
                     paste0(c("Lower confidence bound of hectares (", "Upper confidence bound of hectares ("), confidence, "%, Goodman multinomial) "))


fates <- unique(c(aim_points[["fate"]], lmf_points@data[["fate"]]))

point_counts <- do.call(rbind,
                        lapply(X = split(rbind(aim_points@data[, c("wgtcat", "fate")], lmf_points@data[, c("wgtcat", "fate")]),
                                         rbind(aim_points@data[, c("wgtcat", "fate")], lmf_points@data[, c("wgtcat", "fate")])$wgtcat),
                               fates = fates,
                               FUN = function(X, fates){
                                 output <- data.frame(wgtcat = X[["wgtcat"]][1],
                                                      stringsAsFactors = FALSE)
                                 for (fate in fates) {
                                   output[[paste0("count_", fate)]] <- sum(X[["fate"]] %in% fate)
                                 }
                                 rownames(output) <- NULL
                                 return(output)
                               }))

wgtcat_summary <- merge(x = wgtcat_summary,
                        y = point_counts,
                        all.x = TRUE)

for (fate in fates) {
  wgtcat_summary[[paste0("count_", fate)]][is.na(wgtcat_summary[[paste0("count_", fate)]])] <- 0
}

wgtcat_summary[["in_inference"]] <- !is.na(wgtcat_summary[["sum_relwgts"]])

wgtcat_summary <- wgtcat_summary[, c("wgtcat", "area", "area_units",
                                     "observed_point_count", paste0("count_", fates),
                                     "observed_proportion",
                                     "observed_area", "unobserved_area",
                                     "in_inference")]

names(wgtcat_summary) <- c(wgtcats_var, "area", "area_units",
                           "observed_point_count", paste0("count_", fates),
                           "observed_proportion",
                           "observed_area", "unobserved_area",
                           "in_inference")

wgtcats$wgtcat <-  as.character(wgtcats$wgtcat)
wgtcat_summary <- merge(x = wgtcat_summary,
                        y = wgtcats[, c("design_strata", "wgtcat")],
                        by.x = wgtcats_var,
                        by.y = "wgtcat")

wgtcat_summary <- wgtcat_summary[, c(wgtcats_var, "area", "area_units",
                                     "observed_point_count", paste0("count_", fates),
                                     "observed_proportion",
                                     "observed_area", "unobserved_area",
                                     "in_inference", "design_strata")]

# Add in where they're from
aim_points[["source"]] <- "AIM"
lmf_points[["source"]] <- "LMF"

all_points <- rbind(aim_points[, c("unique_id", "fate", "wgtcat", "source")],
                    lmf_points[, c("unique_id", "fate", "wgtcat", "source")])

all_points <- sp::merge(x = all_points,
                        y = point_weights,
                        all = FALSE)

all_points <- sp::merge(x = all_points,
                        y = ratings[, c("Plot ID", sua_name)],
                        by.x = "unique_id",
                        by.y = "Plot ID",
                        all.x = TRUE)

names(all_points@data) <- c("plot_id", "fate", "wgtcat", "source", "segment", "aim", "lmf", "relwgt", "observed", 
                            "wgt", "reporting_unit", "rating")

all_points <- all_points[, c("plot_id", "rating", "fate", "reporting_unit", "wgtcat", "source", "wgt")]

```

# Reformat for Idaho
```{r}
date <- Sys.Date()

analysis <- select(analysis,
                   "Rating",
                   "Habitat Type",                                                 
 "Indicator",                                                      
 "Number of plots",                                                   
 "Estimated percent of sampled area",                                 
 "Lower confidence bound of percent area (80%, Goodman multinomial) ",
 "Upper confidence bound of percent area (80%, Goodman multinomial) ",
  "Estimated hectares",
 "Lower confidence bound of hectares (80%, Goodman multinomial) ",
"Upper confidence bound of hectares (80%, Goodman multinomial) ")

# reorder ratings
analysis$Rating <-as.factor(analysis$Rating)
levels(analysis$Rating) <-  factor(analysis$Rating)[c(2,1,3)]
analysis <- analysis[order(analysis$Rating),]

# Area totals
# I did this manually already for browns bench but for future analyses:
area_summary <- colSums(x = wgtcat_summary[, c("area","observed_point_count", paste0("count_", fates), "observed_area", "unobserved_area")]) 

area_summary <-  c("Totals:",
                   sum(wgtcat_summary[["area"]]),
                   "NA",
                   colSums(x = wgtcat_summary[, c("observed_point_count", paste0("count_", fates))]),
                   "NA",
                   colSums(x = wgtcat_summary[, c("observed_area", "unobserved_area")]),
                   "total_inference_area" = sum(wgtcat_summary[["area"]][wgtcat_summary[["in_inference"]]==TRUE]),
                   "NA")
area_summary <- as.data.frame(area_summary)
rownames(area_summary) <- colnames(wgtcat_summary)
wgtcat_summary <- rbind(wgtcat_summary,t(area_summary))
```

### Wgtcat .shp
Read in original wgtcat poly, join to wgtcat summary and export
```{r}
wgtcat_spdf <-sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                        layer = "LittleLost_Pahsimeroi_Winter_Strata",
                        stringsAsFactors = FALSE)
## Standardize projection
wgtcat_spdf <- sf::st_transform(wgtcat_spdf,
                                crs = projection)
## Make spatial 
wgtcat_spdf <- methods::as(wgtcat_spdf, "Spatial")

wgtcat_spdf@data <- merge(x = wgtcat_spdf@data,
      y = wgtcat_summary[,c("wgtcat",
                            "area",
                            "area_units",
                            "observed_point_count",
                            paste0("count_",fates),
                            "observed_proportion",
                            "observed_area",
                            "unobserved_area",
                            "in_inference")],
      by.x = wgtcats_var,
      by.y = "wgtcat")

rgdal::writeOGR(wgtcat_spdf,
                dsn = output_path,
                layer = paste0(output_filename,
                               "_wgtcat_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)
```


### Writing Outputs
```{r}
write.csv(point_weights,
          file = paste0(output_path, "/",
                        output_filename,
                        "_pointweights_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(wgtcat_summary,
          file = paste0(output_path, "/",
                        output_filename,
                        "_wgtcatsummary_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(analysis,
          file = paste0(output_path, "/",
                        output_filename,
                        "_analysis_",
                        date, ".csv"),
          row.names = FALSE)
rgdal::writeOGR(all_points,
                dsn = output_path,
                layer = paste0(output_filename,
                               "_points_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)

```
