---
title: "Cox Butte - Winter"
output:
  html_document:
    df_print: paged
---
#### Overview
HAF analysis for the Orijana allotment (a sbsection of the Beatys Butte and Dry Valley Jack Mountain Finescales).


## Orijana Spring

### Setup
Set local library path and load required packages
```{r setup, eval =FALSE}
library(aim.analysis)
library(sp)
library(dplyr)
library(binom)
library(stringr)
```

### Defaults and filepaths

```{r}
# Projection to use. Alber's Equal Area is useful for calculating areas
projection <- sp::CRS("+proj=aea")

# Date of analysis
date <- Sys.Date()

# Confidence level in percent
confidence <- 80

output_path <- "C:\\Users\\alaurencetraynor\\Documents\\2019\\Analysis\\OR\\2020\\Orijana-Cox Butte\\outputs\\April 2020\\Winter"

# Filename for ratings from analysis requestor
ratings_filename <- "HAF_Oregon_2020_CoxButteAllotment_NOCANALYSISREQUEST.xlsx"

# Core filename for all outputs
output_filename <- "Cox_winter"

# The geodatabase and layer names for the spatial data sources
# are embedded in the code below

# Habitat type (corresponds to sheet name in ratings Excel workbook)
# this worksheet has all three SUA ratings so need to filter out unnecessary columns
seasonal_habitat_type = "S-6 Winter"

# The variable in AIM sample designs that contains the unique identifier for each point
aim_idvar <- "PLOT_NM"
# The variable in AIM sample designs that contains the fate for each point
aim_fatevar = "FINAL_DESIG"
# The variable in TerrADat that contains the unique identifier for each point matching AIM designs'
# Will need to rectify odd plot ids from the Dry valley PAC design which do not match Terradat IDs
tdat_idvar <- "PlotID"
# The variable in LMF points that contains the unique identifier for each point
lmf_idvar <- "PLOTKEY"

# The fates in sample designs which correspond to sampled/observed
observed_fates = c("TS", "Target Sampled", "Sampled")

# The fates in sample designs which correspond to unneeded
# (e.g. unused oversample or points designated for future sampling)
invalid_fates = c("NN", NA, "Non Needed", "", "Not Sampled")

# The variable containing the LMF segment identity/membership
# This is found in the points and the segment polygons
segment_var <- "segcode"

# The variable that contains the areas in HECTARES in wgtcats
# This is calculated in ArcMap
wgtcat_area_var = "area_hectares"

# The unique ID for poststratification polygons
wgtcats_var <- "wgtcat"
```

### The filepath to the folder where we have all the spatial data
```{r}
path_spatial <- "C:/Users/alaurencetraynor/Documents/2019/Analysis/OR/2020/Orijana-Cox Butte/inputs/spatial"
```
### The filepath to the folder where we have the Excel stuff
```{r}
path_tabular <- "C:/Users/alaurencetraynor/Documents/2019/Analysis/OR/2020/Orijana-Cox Butte/inputs/tabular"
```

### Read in
#### Reporting unit(s)
Strata and points are already clipped to reporting units in Arc

#### Strata polygons
```{r}
wgtcat_df <-sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                        layer = "wgtcats_cox_winter_final",
                        stringsAsFactors = FALSE)
sf::st_geometry(wgtcat_df) <- NULL
wgtcats <- as.data.frame(wgtcat_df)[,c(1,4,5)]
```

#### Plot Ratings
```{r}
ratings <- readxl::read_xlsx(paste(path_tabular, ratings_filename, sep = "/"),
                             sheet = seasonal_habitat_type)
```

#### Sampled Points 
Make sure points and polygons do not include Z/M dimensions...
```{r}
sampled_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "terradat_winter_cox_final",
                           stringsAsFactors = FALSE)
## Standardize projection
sampled_points <- sf::st_transform(sampled_points,
                                crs = projection)
## Make spatial 
sampled_points <- methods::as(sampled_points, "Spatial")

## sampled plots joined to LMF segments in ArcGIS - POLYID 

## standardise wgtcat variable name
sampled_points@data[["wgtcat"]] <- sampled_points@data[[wgtcats_var]]

# LMF points:
lmf_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "lmf_winter_cox_final",
                           stringsAsFactors = FALSE)
## Standardize projection
lmf_points <- sf::st_transform(lmf_points,
                                crs = projection)
## Make spatial 
lmf_points <- methods::as(lmf_points, "Spatial")
```

#### Design points
```{r}
### All design points were merge in ArcMap
design_points <- sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                           layer = "design_points_cox_winter_final",
                           stringsAsFactors = FALSE)
## Standardize projection
design_points <- sf::st_transform(design_points,
                                crs = projection)
## Make spatial 
design_points <- methods::as(design_points, "Spatial")
```

###### Sort out points ######
#### Prepare TerrADat points
```{r}
## Fix Dry Valley PAC point IDs by joining to design points with Plot Key
sampled_points <- sp::merge(x = sampled_points,
                        y = design_points@data[,c(6,7)], 
                        by.x = "PlotKey", 
                        by.y = "PLOT_KEY")

# Also add these full plot IDs to the ratings spreadsheet
ratings <- dplyr::right_join(x = sampled_points@data[,c("PlotID", "PLOT_NM")],
                 y = ratings,
                 by = c("PlotID"="Plot Identifier"))

### Make naming consistent across all data
# Move unique IDs to unique_id in TerrADat
sampled_points@data[["unique_id"]] <- sampled_points@data[["PLOT_NM"]]

# Set fate variable in TerrADat
sampled_points@data[["fate"]] <- observed_fates[1]

# Move LMF segment membership to segment
sampled_points@data[["segment"]] <- sampled_points@data[[segment_var]]

# repeat for all points
design_points@data[["unique_id"]] <- design_points@data[[aim_idvar]]
design_points@data[["fate"]] <- design_points@data[[aim_fatevar]]
design_points@data[["wgtcat"]] <- design_points@data[[wgtcats_var]]
design_points@data[["segment"]] <- design_points@data[[segment_var]]

lmf_points@data[["unique_id"]] <- lmf_points@data[[lmf_idvar]]
lmf_points@data[["fate"]] <- observed_fates[1]
lmf_points@data[["wgtcat"]] <- lmf_points@data[[wgtcats_var]]
lmf_points@data[["segment"]] <- lmf_points@data[[segment_var]]

lmf_points <- lmf_points[, c("unique_id", "fate", "wgtcat", "segment")]
```


#### Create combined points shapefile
### Combine all sample design points
```{r}
### Mash the design files together, keeping only necessary variables
design_points <- design_points[, c("unique_id", "fate", "wgtcat", "segment")] 

### Check all terradatsampled plots are in design file and vice versa 
sum(sampled_points$unique_id %in% design_points$unique_id)
sum(design_points$unique_id[design_points$fate %in% observed_fates] %in% sampled_points$unique_id)
length(sampled_points$unique_id) # one extra sampled point "Rock Creek Ranch" - potenitally targeted not rated for HAF - lets remove it

sampled_points <- sampled_points[sampled_points[["unique_id"]]!="Rock Creek Ranch"]
```
### QC
```{r}
# Checking against the ratings confirms if all the rated plots are represented (important)
# and if all the sampled_plots are rated (not important, but informative)
if (!all(ratings[["PLOT_NM"]][ratings[["AIM"]] == "AIM"] %in% sampled_points$PLOT_NM)) {
  warning("NOT ALL RATED AIM PLOTS ARE PRESENT IN sampled_points!")
}
# Ratings sheet uses the Terradat plot IDs
all(sampled_points@data$PLOT_NM %in% ratings[["PLOT_NM"]]) # several AIM plots out of seasonal dates 

# filter sampled plots by those that were rated:
sampled_points <- sampled_points[sampled_points[["PLOT_NM"]] %in% ratings[["PLOT_NM"]], ] 
lmf_points <- lmf_points[lmf_points[["unique_id"]] %in% ratings[["PlotID"]], ]
```

# Now combine:
# All design_points which were not observed. This removes any points that overlap with sampled_points but also
# any that were observed but somehow not in TerrADat and deemed unsuitable for rating
# which therefore means they also should not be included in weighting
# All sampled points (already restricted to the ones which were rated)

```{r}
aim_points <- rbind(design_points[!(design_points$fate %in% observed_fates), c("unique_id", "fate", "wgtcat", "segment")], 
                    sampled_points[, c("unique_id", "fate", "wgtcat", "segment")])

# Take out invalid fates, e.g. things that don't get included in weighting
aim_points <- aim_points[!(aim_points$fate %in% invalid_fates), ] # no rejected points in design

# Add in coordinates (needed for the analysis step)
aim_points <-  aim.analysis::add_coords(aim_points,
                                        xynames = c("x", "y"))

lmf_points <-  aim.analysis::add_coords(lmf_points,
                                        xynames = c("x", "y"))

aim_points$wgtcat <- as.character(aim_points$wgtcat) 
lmf_points$wgtcat <- as.character(lmf_points$wgtcat) 
```

###### Analyze ######
############################################################
#### Calculate weights ####
### Note that this is the contents of aim.analysis::weight_aimlmf() verbatim
### As a function, it was producing errors that it does not as bare code
############################################################

This functions is giving errors:
aim.analysis::weight_aimlmf(aim_points = aim_points,
                            lmf_points = lmf_points,
                            aim_idvar <- "unique_id",
                            lmf_idvar <- "unique_id",
                            aim_fatevar <- "fate",
                            wgtcat_var <- "wgtcat",
                            segments <- NULL,
                            segment_var <- "segment",
                            verbose <- FALSE )
Using raw code instead:                     
```{r}
# These are being redefined here as though they were arguments to the function
aim_idvar <- "unique_id"
lmf_idvar <- "unique_id"
aim_fatevar <- "fate"
wgtcat_var <- "wgtcat"
segments <- NULL
segment_var <- "segment"
verbose <- FALSE

if (length(wgtcat_var) != 1 | class(wgtcat_var) != "character") {
  stop("wgtcat_var must be a single character string")
}
if (!is.null(wgtcat_area_var)) {
  if (length(wgtcat_area_var) != 1 | class(wgtcat_area_var) != "character") {
    stop("wgtcat_area_var must be a single character string")
  }
}
if (length(segment_var) != 1 | class(segment_var) != "character") {
  stop("segment_var must be a single character string")
}

if (!is.null(wgtcats)) {
  if (!(class(wgtcats) %in% c("SpatialPolygonsDataFrame", "data.frame"))) {
    stop("wgtcats must be a spatial polygons data frame or data frame")
  }
  if (nrow(wgtcats) < 1) {
    stop("wgtcats contains no observations/data")
  }
  if (!(wgtcat_var %in% names(wgtcats))) {
    stop(paste("The variable", wgtcat_var, "does not appear in wgtcats@data"))
  }
}

if (!is.null(segments)) {
  if (!(class(segments) %in% "SpatialPolygonsDataFrame")) {
    stop("segments must be a spatial polygons data frame")
  }
  if (nrow(segments) < 1) {
    stop("segments contains no observations/data")
  }
  if (!(segment_var %in% names(segments))) {
    stop(paste("The variable", segment_var, "does not appear in segments@data"))
  }
}

if (is.null(aim_fatevar)) {
  warning("No fate variable specified for AIM points. Assuming all were observed/sampled.")
  aim_fatevar <- "fate"
  aim_points[["fate"]] <- "observed"
  lmf_points[["fate"]] <- "observed"
  observed_fates <- "observed"
} else {
  if (length(aim_fatevar) > 1 | class(aim_fatevar) != "character") {
    stop("The aim fate variable must be a single character string")
  }
  if (!aim_fatevar %in% names(aim_points)) {
    stop(paste("The variable", aim_fatevar, "does not appear in aim_points@data"))
  } else {
    aim_points[["fate"]] <- aim_points[[aim_fatevar]]
  }
  if (is.null(observed_fates)) {
    warning("No observed fates provided. Assuming all AIM points were observed/sampled unless specified otherwise with invalid_fates")
    observed_fates <- unique(aim_points[["fate"]])
    observed_fates <- observed_fates[!(observed_fates %in% invalid_fates)]
  }
}

lmf_points[["fate"]] <- observed_fates[1]

if (!is.null(observed_fates)) {
  if (!any(aim_points[["fate"]] %in% observed_fates)) {
    warning("No AIM points have a fate specified as observed.")
  }
}

# Harmonize projections
if (is.null(projection)) {
  if (!is.null(wgtcats)) {
    projection <- wgtcats@proj4string
  } else if (!is.null(segments)) {
    projection <- segments@proj4string
  }
}

if (!is.null(projection)) {
  if (class(aim_points) %in% c("SpatialPointsDataFrame")) {
    if (!identical(projection, aim_points@proj4string)) {
      aim_points <- sp::spTransform(aim_points,
                                    CRSobj = projection)
    }
  }
  if (class(lmf_points) %in% c("SpatialPointsDataFrame")) {
    if (!identical(projection, lmf_points@proj4string)) {
      lmf_points <- sp::spTransform(lmf_points,
                                    CRSobj = projection)
    }
  }
  if (!is.null(wgtcats)) {
    if (class(lmf_points) %in% c("SpatialPolygonsDataFrame")) {
      if (!identical(projection, wgtcats)) {
        wgtcats <- sp::spTransform(wgtcats,
                                   CRSobj = projection)
      }
    }
  }
  if (!is.null(segments)) {
    if (!identical(projection, segments@proj4string)) {
      segments <- sp::spTransform(segments,
                                  CRSobj = projection)
    }
  }
}



# Assign the weight categories
if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
  if (!(wgtcat_var %in% names(wgtcats))) {
    stop("The variable ", wgtcat_var, " does not appear in wgtcats")
  }
  aim_points[["wgtcat"]] <- sp::over(aim_points, wgtcats)[[wgtcat_var]]
  lmf_points[["wgtcat"]] <- sp::over(lmf_points, wgtcats)[[wgtcat_var]]
} else {
  if (!(wgtcat_var %in% names(aim_points))) {
    stop("The variable ", wgtcat_var, " does not appear in aim_points")
  }
  if (!(wgtcat_var %in% names(lmf_points))) {
    stop("The variable ", wgtcat_var, " does not appear in lmf_points")
  }
  aim_points[["wgtcat"]] <- aim_points[[wgtcat_var]]
  lmf_points[["wgtcat"]] <- lmf_points[[wgtcat_var]]
}


# Assign the LMF segment codes
if (is.null(segments)) {
  if (!(segment_var %in% names(aim_points))) {
    stop("The variable ", segment_var, " does not appear in aim_points")
  }
  if (!(segment_var %in% names(lmf_points))) {
    stop("The variable ", segment_var, " does not appear in lmf_points")
  }
  aim_points[["segment"]] <- aim_points[[segment_var]]
  lmf_points[["segment"]] <- lmf_points[[segment_var]]
} else {
  if (!(segment_var %in% names(segments))) {
    stop("The variable ", segment_var, " does not appear in segments")
  }
  aim_points[["segment"]] <- sp::over(aim_points, segments)[[segment_var]]
  lmf_points[["segment"]] <- sp::over(lmf_points, segments)[[segment_var]]
}


# Just harmonize the idvar names for now
aim_points[["unique_id"]] <- aim_points[[aim_idvar]]
lmf_points[["unique_id"]] <- lmf_points[[lmf_idvar]]

# Add reporting units
aim_points[["reporting_unit"]] <- seasonal_habitat_type
lmf_points[["reporting_unit"]] <- seasonal_habitat_type

# If somehow LMF points aren't in a segment, that's a major problem
if (any(is.na(lmf_points[["segment"]]))) {
  stop(paste("The following LMF points did not spatially intersect any segment polygons:",
             paste(lmf_points[is.na(lmf_points[["segment"]]), "unique_id"], collapse = ", ")))
}

# TODO: Stick a check in here that the segment ID from the polygons
# matches the one derived from the LMF plot ID
# Probably just warn if not?


# Get data frames
if (class(aim_points) %in% "SpatialPointsDataFrame") {
  aim_df <- aim_points@data
} else {
  aim_df <- aim_points
}
if (class(lmf_points) %in% "SpatialPointsDataFrame") {
  lmf_df <- lmf_points@data
} else {
  lmf_df <- lmf_points
}

# NOTE THAT THIS FILTERS OUT ANYTHING FLAGGED AS NOT NEEDED IN THE FATE
# So that'd be unused oversamples or points from the FUTURE that no one would've sampled anyway
aim_df <- aim_df[!(aim_df[["fate"]] %in% invalid_fates), c("unique_id", "fate", "wgtcat", "segment")]
aim_df[["aim"]] <- TRUE
aim_df[["lmf"]] <- FALSE

# We only have target sampled LMF points available to us, so we don't need to filter them
lmf_df <- lmf_df[, c("unique_id", "fate", "wgtcat", "segment")]
lmf_df[["aim"]] <- FALSE
lmf_df[["lmf"]] <- TRUE

# Combine them
combined_df <- unique(rbind(aim_df, lmf_df))

# There shouldn't be any that don't belong to a wgtcat anymore
combined_df <- combined_df[!is.na(combined_df[["wgtcat"]]), ]

# Add an observed variable for easy reference later
combined_df[["observed"]] <- combined_df[["fate"]] %in% observed_fates

# To make the lookup table, drop any points that fell outside LMF segments
combined_segmentsonly_df <- combined_df[!is.na(combined_df[["segment"]]), ]

# Create a segment relative weight lookup table
segment_relwgt_lut <- do.call(rbind,
                              lapply(X = split(combined_segmentsonly_df, combined_segmentsonly_df[["segment"]]),
                                     FUN = function(X){
                                       # These are the count of AIM points with any valid fate
                                       aim_count <- sum(X[["aim"]])
                                       
                                       # We also need the count of LMF points with any valid fate, but that's complicated
                                       # We only have the sampled LMF points so we can count those
                                       lmf_sampled_count <- sum(X[["lmf"]])
                                       # To get the number of evaluated but not sampled points:
                                       # The LMF plot keys end in a digit that represents the intended sampling order within a segment
                                       # 1 and 2 are considered base points and were intended to be sampled
                                       # If a sampled LMF plot's plot key ends in 3, that means that one or both of the base points
                                       # were evaluated and rejected rather than sampled, which brings the evaluated LMF plot count
                                       # to three for the segment.
                                       # This just asks if the third point was used
                                       lmf_oversample_used <- any(grepl(X[["unique_id"]][X[["lmf"]]],
                                                                        pattern = "\\D3$"))
                                       
                                       # Likewise, if only one LMF plot was sampled in a segment, that means the other two were
                                       # evalurated and rejected rather than sampled, also bringing the total to three.
                                       # So if there was only one sampled or if the oversample was used, there were three evaluated
                                       if (sum(X[["lmf"]]) == 1 | lmf_oversample_used) {
                                         lmf_count <- 3
                                       } else {
                                         # This will fire only if there sampled count was 2, but better to be safe here
                                         lmf_count <- lmf_sampled_count
                                       }
                                       
                                       
                                       # The relative weight for points falling within a segment is calculated as
                                       # 1 / (number of points)
                                       relative_weight <- 1 / sum(aim_count, lmf_count)
                                       
                                       output <- data.frame("segment" = X[["segment"]][1],
                                                            "relwgt" = relative_weight,
                                                            stringsAsFactors = FALSE)
                                       return(output)
                                     }))

# Add the relative weights to the combined AIM and LMF points
combined_df <- merge(x = combined_df,
                     y = segment_relwgt_lut,
                     by = "segment",
                     all.x = TRUE)

# manualy fix rel weight of segment that was cut by seasonal habitst area
combined_df$relwgt[combined_df$segment=="24"] <- 2/3

# Anywhere there's an NA associated with an AIM point, that's just one that fell outside the segments
combined_df[is.na(combined_df[["relwgt"]]) & combined_df[["aim"]], "relwgt"] <- 1



if (is.null(wgtcat_area_var)) {
  if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
    wgtcat_df <- aim.analysis::add.area(wgtcats)@data
  } else {
    stop("No name for a variable in wgtcats containing the area in hectares was provided and wgtcats is not a spatial polygons data frame so area can't be calculated")
  }
} else {
  if (!(wgtcat_area_var %in% names(wgtcats))) {
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- aim.analysis::add.area(wgtcats)@data
    } else {
      stop("The variable ", wgtcat_area_var, " does not appear in wgtcats")
    }
  } else {
    warning("Trusting that the variable ", wgtcat_area_var, " in wgtcats contains the areas in hectares")
    if (class(wgtcats) %in% "SpatialPolygonsDataFrame") {
      wgtcat_df <- wgtcats@data
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    } else {
      wgtcat_df <- wgtcats
      wgtcat_df[["AREA.HA"]] <- wgtcat_df[[wgtcat_area_var]]
    }
  }
}


wgtcat_df[["wgtcat"]] <- wgtcat_df[[wgtcat_var]]

# Making sure that these aren't spread out over multiple observations
if (verbose) {
  message("Summarizing wgtcat_df by wgtcat to calculate areas in case the wgtcats are split into multiple observations")
}
wgtcat_df <- dplyr::summarize(dplyr::group_by(wgtcat_df,
                                              wgtcat),
                              "hectares" = sum(AREA.HA))

wgtcat_areas <- setNames(wgtcat_df[["hectares"]], wgtcat_df[["wgtcat"]])

# had a issue here reading wgtcats as integers
wgtcat_df$wgtcat <-  as.character(wgtcat_df$wgtcat)
  
weight_info <- lapply(X = wgtcat_df[["wgtcat"]],
                      wgtcat_areas = wgtcat_areas,
                      points = combined_df,
                      wgtcat_var = wgtcat_var,
                      FUN = function(X, wgtcat_areas, points, wgtcat_var){
                        # Area of this polygon
                        area <- wgtcat_areas[X]
                        
                        # All of the points (AIM and LMF) falling in this polygon
                        points <- points[points[["wgtcat"]] == X, ]
                        
                        # If there are in fact points, do some weight calculations!
                        if (nrow(points) > 0 & any(points[["observed"]])) {
                          # The number of observed AIM points with a relative weight of 1
                          # So, not sharing an LMF segment with any LMF points
                          aim_standalone <- sum(combined_df[["aim"]] & combined_df[["relwgt"]] == 1)
                          # The number of unique segments selected for LMF points
                          lmf_segment_count <- length(unique(points[["segment"]]))
                          
                          # The approximate number of 160 acre segments in this polygon
                          # Obviously, not all of the segments were selected in the first stage
                          # of the LMF design, but this is how many were available in this polygon
                          approximate_segment_count <- area / (160 / 2.47)
                          
                          # This is the sum of the relative weights of the OBSERVED points
                          # This does not include the inaccessible, rejected, or unknown points
                          # It does include both AIM and LMF, however
                          sum_observed_relwgts <- sum(points[points[["observed"]], "relwgt"])
                          # This is the sum of the relative weights of all the AIM points, regardless of fate
                          sum_relwgts <- sum(points[["relwgt"]])
                          
                          # The units are segments per point
                          # The segments are 120 acre quads (quarter sections?) that the first stage of LMF picked from
                          # Steve Garman called this "ConWgt" which I've expanded to conditional_weight
                          # but that's just a guess at what "con" was short for
                          conditional_weight <- approximate_segment_count / (aim_standalone + lmf_segment_count)
                          conditional_area <- conditional_weight * sum_observed_relwgts
                          
                          # What's the observed proportion of the area?
                          observed_proportion <- sum_observed_relwgts / sum_relwgts
                          # And how many acres is that then?
                          # We can derive the "unknown" or "unsampled" area as the difference
                          # between the polygon area and the observed area
                          observed_area <- area * observed_proportion
                          
                          # Then this adjustment value is calculated
                          weight_adjustment <- observed_area / conditional_area
                          
                          
                          
                          # Put everything about the wgtcat in general in one output data frame
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = approximate_segment_count,
                                                      sum_observed_relwgts = sum_observed_relwgts,
                                                      sum_relwgts = sum_relwgts,
                                                      observed_proportion = observed_proportion,
                                                      observed_area = observed_area,
                                                      unobserved_area = area - observed_area,
                                                      conditional_weight = conditional_weight,
                                                      conditional_area = conditional_area,
                                                      weight_adjustment = weight_adjustment,
                                                      point_count = sum(points[["observed"]]),
                                                      observed_point_count = nrow(points),
                                                      stringsAsFactors = FALSE)
                          
                          # But much more importantly add the calculated weights to the points
                          output_points <- points
                          
                          # This handles situations where there were points, but none of them were observed
                          # In that case, weight_adjustment will be 0 / 0 = NaN
                          # That's fine because we can identify that this polygon is still in the inference area
                          # but that its entire area is "unknown" because no data were in it
                          if (is.nan(weight_adjustment)) {
                            output_points[["wgt"]] <- NA
                          } else {
                            # The point weight is calculated here.
                            # This is Garman's formula and I don't have the documentation justifying it on hand
                            output_points[["wgt"]] <- weight_adjustment * conditional_weight * points[["relwgt"]]
                            message("Checking weight sum for ", X)
                            # I'm rounding here because at unrealistically high levels of precision it gets weird and can give false positives
                            if (round(sum(output_points[["wgt"]]), digits = 3) != round(area, digits = 3)) {
                              warning("The sum of the point weights (", sum(output_points[["wgt"]]), ") does not equal the polygon area (", area, ") for ", X)
                            }
                          }
                          
                        } else {
                          # Basically just empty data frames
                          output_wgtcat <- data.frame(wgtcat = X,
                                                      area = area,
                                                      area_units = "hectares",
                                                      approximate_segment_count = area / (160 / 2.47),
                                                      sum_observed_relwgts = NA,
                                                      sum_relwgts = NA,
                                                      observed_proportion = 0,
                                                      observed_area = 0,
                                                      unobserved_area = area,
                                                      conditional_weight = NA,
                                                      conditional_area = NA,
                                                      weight_adjustment = NA,
                                                      point_count = 0,
                                                      observed_point_count = 0,
                                                      stringsAsFactors = FALSE)
                          output_points <- NULL
                        }
                        
                        return(list(points = output_points,
                                    wgtcat = output_wgtcat))
                      })

point_weights <- do.call(rbind,
                         lapply(X = weight_info,
                                FUN = function(X){
                                  X[["points"]]
                                }))
wgtcat_summary <- do.call(rbind,
                          lapply(X = weight_info,
                                 FUN = function(X){
                                   X[["wgtcat"]]
                                 }))

```
# Check LMF segment 24 for 2 rejections

###### Analysis ######
```{r}
all_points <- rbind(aim_points[, c("unique_id", "x", "y")],
                    lmf_points[, c("unique_id", "x", "y")])

ratings$PLOT_NM <- ifelse(is.na(ratings$PLOT_NM),
                         ratings$PlotID,
                          ratings$PLOT_NM)
  
# Add the ratings to the point info and convert to a data frame
all_points <- merge(x = all_points@data,
                    y = ratings,
                    by.x = "unique_id",
                    by.y = "PLOT_NM",
                    all.x = FALSE)

# Add in the indicator, which for these will always be "Habitat suitability"
all_points[["indicator"]] <- "Habitat suitability"

# Add in the season (this is defined up above) e.g. "S-4 Upland Summer Late Brood-rearing Habitat"
all_points[["reporting_unit"]] <- "Cox_Winter"
point_weights[["reporting_unit"]] <- "Cox_Winter"

# And analyze
analysis <- aim.analysis::analyze(benchmarked_points = all_points,
                                  point_weights = point_weights,
                                  id_var = "unique_id",
                                  indicator_var = "indicator",
                                  value_var = "Suitability",
                                  x_var = "x",
                                  y_var = "y",
                                  reporting_var = "reporting_unit",
                                  weight_var = "wgt",
                                  conf = confidence)
```

### Confidence Interval Function
# 

```{r}
goodman_cis <- function(counts,
                        alpha = 0.2,
                        chisq = "best",
                        verbose = FALSE){
  
  if (!(chisq %in% c("A", "B", "best"))) {
    stop("The only valid values for chisq are 'A', 'B', and 'best'.")
  }
  
  # Goodman describes the upper and lower bounds with the equations:
  # Lower estimated pi_i = {A + 2n_i - {A[A + 4n_i(N - n_i) / N]}^0.5} / [2(N + A)]
  # Upper estimated pi_i = {A + 2n_i + {A[A + 4n_i(N - n_i) / N]}^0.5} / [2(N + A)]
  
  # n_i is the "observed cell frequencies in population of size N" (aka count of observations) from a category
  # so that's the incoming argument counts. We'll rename for consistency with the original math (and statistics as a discipline)
  n <- counts
  
  # N is the population those counts make up, or, in lay terms, the total observation count
  N <- sum(counts)
  
  # k is the number of categories the population has been sorted into
  # Useful for degrees of freedom
  k <- length(counts)
  
  # "A is the upper alpha * 100-th percentage point of the chi-square distribution with k - 1 degrees of freedom"
  # and B is an alternative which uses alpha / k and one degree of freedom
  # Goodman states that B should be less than A for situations
  # where k > 2 AND alpha is 0.1, 0.05, or 0.01.
  chisq_quantiles <- c("A" = stats::qchisq(p = 1 - alpha,
                                           df = k - 1),
                       "B" = stats::qchisq(p = 1 - (alpha / k),
                                           df = 1))
  
  
  # According to Goodman, A and B are both valid options for the chi-square quantile
  # So the user can specify which they want or just ask for the one that minimizes the confidence intervals
  chisq_quantile <- switch(chisq,
                           "A" = {chisq_quantiles["A"]},
                           "B" = {chisq_quantiles["B"]},
                           "best" = {
                             pick <- which.min(chisq_quantiles)
                             if (verbose){
                               switch(names(chisq_quantiles)[pick],
                                      "A" = message("The chi-square quantile calculation that will provide the tighter confidence intervals is A, the upper alpha X 100-th percentage point of the chi-square distribution with k - 1 degrees of freedom"),
                                      "B" = message("The chi-square quantile calculation that will provide the tighter confidence intervals is B, the upper alpha / k X 100-th percentage point of the chi-square distribution with 1 degree of freedom"))
                             }
                             chisq_quantiles[pick]
                           })
  
  # Calculate the bounds!
  # Note that these ARE symmetrical, just not around the proportions.
  # They're symmetrical around chisq_quantile + 2 * n / (2 * (N + chisq_quantile))
  # The variable A has been replaced with chisq_quantile because it may be A or B, depending
  # Since the only multi-value vector involved here is n, these will be vectors of length k,
  # having one value for each of the values in n and in the same order as n
  lower_bounds <- (chisq_quantile + 2 * n - sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))) / (2 * (N + chisq_quantile))
  upper_bounds <- (chisq_quantile + 2 * n + sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))) / (2 * (N + chisq_quantile))
  
  # A proportion can never be greater than 1 or less than 0 (duh)
  # That's definitely a thing that can happen if the magnitude of sqrt(chisq_quantile * (chisq_quantile + 4 * n * (N - n) / N))
  # is large enough
  lower_bounds[lower_bounds < 0] <- 0
  upper_bounds[upper_bounds > 1] <- 1
  
  # Build the output
  output <- data.frame(count = n,
                       proportion = n / N,
                       lower_bound = lower_bounds,
                       upper_bound = upper_bounds)
  
  # What are the categories called? If anything, that is
  k_names <- names(n)
  
  if (!is.null(k_names)) {
    output[["category"]] <- k_names
    output <- output[, c("category", "count", "proportion", "lower_bound", "upper_bound")]
  }
  
  return(output)
}

```

### Formatting Outputs
# don't need Goodman's multinomial as there are only 2 categories present - can stick with Wilson binomial caluclated above
```{r}
analysis <- analysis[, c("Subpopulation",
                         "Indicator",
                         "Category",
                         "NResp",
                         "Estimate.P",
                         "Estimate.U")]

adjusted_counts <- (sum(analysis[["NResp"]][analysis[["Category"]] != "Total"]) * analysis[["Estimate.P"]][analysis[["Category"]] != "Total"] / 100)

# Add in thrid category with 0 counts
adjusted_counts <- c(adjusted_counts,0)

goodman_cis <- goodman_cis(counts = adjusted_counts)

analysis <- rbind.data.frame(analysis, c("Cox_Winter", "Habitat suitability","Unsuitable",0,NA,NA))
                  
goodman_cis[["Category"]] <- analysis[["Category"]][analysis[["Category"]] != "Total"]

analysis <- merge(x = analysis,
                  y = goodman_cis[, c("Category", "lower_bound", "upper_bound")])

# estimate is converted to character somewhere up here :(
# Calculate the hectares using the Wilson binomial confidence intervals from percentage
total_hectares <- as.numeric(analysis[1,"Estimate.U"])/(as.numeric(analysis[1,"Estimate.P"])/100)
  
analysis[["LCB80_area"]] <- analysis[["lower_bound"]]  * total_hectares
analysis[["UCB80_area"]] <- analysis[["upper_bound"]]  * total_hectares

analysis[["lower_bound"]] <- analysis[["lower_bound"]]*100
analysis[["upper_bound"]] <- analysis[["upper_bound"]]*100

names(analysis) <- c("Rating","Reporting Unit",
                     "Indicator",
                     "Number of plots",
                     "Estimated percent of sampled area",
                     "Estimated hectares",
                     paste0(c("Lower confidence bound of percent area (", "Upper confidence bound of percent area ("), confidence, "%, Goodman's Multinomial) "),
                     paste0(c("Lower confidence bound of hectares (", "Upper confidence bound of hectares ("), confidence, "%, Goodman's Multinomial) "))


fates <- unique(c(aim_points[["fate"]], lmf_points@data[["fate"]]))

point_counts <- do.call(rbind,
                        lapply(X = split(rbind(aim_points@data[, c("wgtcat", "fate")], lmf_points@data[, c("wgtcat", "fate")]),
                                         rbind(aim_points@data[, c("wgtcat", "fate")], lmf_points@data[, c("wgtcat", "fate")])$wgtcat),
                               fates = fates,
                               FUN = function(X, fates){
                                 output <- data.frame(wgtcat = X[["wgtcat"]][1],
                                                      stringsAsFactors = FALSE)
                                 for (fate in fates) {
                                   output[[paste0("count_", fate)]] <- sum(X[["fate"]] %in% fate)
                                 }
                                 rownames(output) <- NULL
                                 return(output)
                               }))

wgtcat_summary <- merge(x = wgtcat_summary,
                        y = point_counts,
                        all.x = TRUE)

for (fate in fates) {
  wgtcat_summary[[paste0("count_", fate)]][is.na(wgtcat_summary[[paste0("count_", fate)]])] <- 0
}

wgtcat_summary[["in_inference"]] <- !is.na(wgtcat_summary[["sum_relwgts"]])

wgtcat_summary <- wgtcat_summary[, c("wgtcat", "area", "area_units",
                                     "observed_point_count", paste0("count_", fates),
                                     "observed_proportion",
                                     "observed_area", "unobserved_area",
                                     "in_inference")]

names(wgtcat_summary) <- c(wgtcats_var, "area", "area_units",
                           "observed_point_count", paste0("count_", fates),
                           "observed_proportion",
                           "observed_area", "unobserved_area",
                           "in_inference")

wgtcats$wgtcat <-  as.character(wgtcats$wgtcat)
wgtcat_summary <- merge(x = wgtcat_summary,
                        y = wgtcats[, c("strata_concat", "wgtcat")],
                        by.x = wgtcats_var,
                        by.y = "wgtcat")

wgtcat_summary <- wgtcat_summary[, c(wgtcats_var, "area", "area_units",
                                     "observed_point_count", paste0("count_", fates),
                                     "observed_proportion",
                                     "observed_area", "unobserved_area",
                                     "in_inference", "strata_concat")]

# Add in where they're from
aim_points[["source"]] <- "AIM"
lmf_points[["source"]] <-  "LMF"

all_points <- rbind(aim_points[, c("unique_id", "fate", "wgtcat", "source")],
                    lmf_points[, c("unique_id", "fate", "wgtcat", "source")])

all_points <- sp::merge(x = all_points,
                        y = point_weights,
                        all = FALSE)

all_points <- sp::merge(x = all_points,
                        y = ratings[, c("PLOT_NM", "Suitability")],
                        by.x = "unique_id",
                        by.y = "PLOT_NM",
                        all.x = TRUE)

names(all_points@data) <- c("plot_id", "fate", "wgtcat", "source", "segment", "aim", "lmf", "relwgt", "observed", 
                            "wgt", "reporting_unit", "rating")

all_points <- all_points[, c("plot_id", "rating", "fate", "reporting_unit", "wgtcat", "source", "wgt")]

```

# Reformat to make output pretty
```{r}
date <- Sys.Date()

analysis <- select(analysis,
                   "Rating",
                   "Reporting Unit",                                                 
 "Indicator",                                                      
 "Number of plots",                                                   
 "Estimated percent of sampled area",                                 
 "Lower confidence bound of percent area (80%, Goodman's Multinomial) ",
 "Upper confidence bound of percent area (80%, Goodman's Multinomial) ",
  "Estimated hectares",
 "Lower confidence bound of hectares (80%, Goodman's Multinomial) ",
"Upper confidence bound of hectares (80%, Goodman's Multinomial) ")

# reorder ratings
analysis$Rating <-as.factor(analysis$Rating)
levels(analysis$Rating) <-  factor(analysis$Rating)[c(2,1,3)]
analysis <- analysis[order(analysis$Rating),]

# Area totals
# I did this manually already for browns bench but for future analyses:
area_summary <- colSums(x = wgtcat_summary[, c("area","observed_point_count", paste0("count_", fate), "observed_area", "unobserved_area")]) 

area_summary <-  c("Totals:",
                   sum(wgtcat_summary[["area"]]),
                   "NA",
                   colSums(x = wgtcat_summary[, c("observed_point_count", paste0("count_", fate))]),
                   "NA",
                   colSums(x = wgtcat_summary[, c("observed_area", "unobserved_area")]),
                   "total_inference_area" = sum(wgtcat_summary[["area"]][wgtcat_summary[["in_inference"]]==TRUE]),
                   "NA")
area_summary <- as.data.frame(area_summary)
rownames(area_summary) <- colnames(wgtcat_summary)
wgtcat_summary <- rbind(wgtcat_summary,t(area_summary))
```

### Wgtcat .shp
Read in original wgtcat poly, join to wgtcat summary and export
```{r}
wgtcat_spdf <-sf::st_read(dsn = paste(path_spatial, "SampleFrameCreation.gdb", sep = "/"),
                        layer = "wgtcats_cox_winter_final",
                        stringsAsFactors = FALSE)
## Standardize projection
wgtcat_spdf <- sf::st_transform(wgtcat_spdf,
                                crs = projection)

## Giving errors due to "MULTISURFACE" features - so simplified these to "MULTIPOLYGONS"
wgtcat_spdf_simplified <- sf::st_cast(wgtcat_spdf, to = "MULTIPOLYGON")

## Make spatial 
wgtcat_spdf <- methods::as(wgtcat_spdf, "Spatial")

wgtcat_spdf <- sp::merge(x = wgtcat_spdf,
      y = wgtcat_summary[,c(1:9)],
      by.x = wgtcats_var,
      by.y = "wgtcat")

rgdal::writeOGR(wgtcat_spdf,
              dsn = output_path,
               layer = paste0(output_filename,
                               "_wgtcat_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)
```
### Writing Outputs
```{r}
write.csv(point_weights,
          file = paste0(output_path, "/",
                        output_filename,
                        "_pointweights_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(wgtcat_summary,
          file = paste0(output_path, "/",
                        output_filename,
                        "_wgtcatsummary_",
                        date, ".csv"),
          row.names = FALSE)
write.csv(analysis,
          file = paste0(output_path, "/",
                        output_filename,
                        "_analysis_",
                        date, ".csv"),
          row.names = FALSE)

 rgdal::writeOGR(all_points,
                dsn = output_path,
                layer = paste0(output_filename,
                               "_points_",
                               date),
                driver = "ESRI Shapefile",
                overwrite_layer = TRUE)

```
