---
title: "Brown Bench Site Scale HAF Weighted Analysis"
output:
  html_document:
    df_print: paged
---
#### Overview
- Proportional area estimates for 3 different reporting units:
  1. Each Seasonal Use Area (SUA) within the Brown's Bench fine-scale area
  2. Allotment group 1
  3. Allotment group 2
- Include AIM, LMF, and modified HAF plots 
- Sampling designs include: 
  1. LMF - track down original sample design - 
  2. Two modified HAF shiny tool designs (2017 & 2018) - looks good
  3. Elko: Track these down 
    - 2011 - \ProjectsNational\AIM\Projects\NV\State\Data\Sample_Design\AIM_SDD_NVAIM_20160106.gdb
    - 2015 - \ProjectsNational\AIM\Projects\NV\State\Data\Sample_Design\AIM_SDD_NVAIM_20160106.gdb 
    - 2017 (actually 2016 design) - (not strata?)\ProjectsNational\AIM\Projects\NV\State\Data\Sample_Design\2016\NV_StateOffice_2016-2020_WORKINGCOPY.gdb
  4. Burley FO AIM 2016 - \ProjectsNational\AIM\Projects\ID\Burley_FO\Data\Sample Design\2016\
  5. Jarbidge 2017 - \ProjectsNational\AIM\Projects\ID\Jarbridge\Data\SampleDesign\2017\ID_JarbridgeFO_2017-2021_SDD.gdb
  6. ID state ESD ??? - these are basically targeted points so leave out of weighted analysis
- Area includes the Burley (ID), Jarbidge (ID) and Elko (NV) field offices


#### SG code examples from Anthony:
Parent folder:  T:\ProjectsNational\GRSG\HAF\Assessments_archived_032019\OR_2017_Warners\Sitescale
Code blocks:  T:\ProjectsNational\GRSG\HAF\Assessments_archived_032019\OR_2017_Warners\Sitescale\Methods\Code
I think the heavy lifter is the proportions.r file
Outputs we provided back to the field:  T:\ProjectsNational\GRSG\HAF\Assessments_archived_032019\OR_2017_Warners\Sitescale\Results\

#### Setup
Set local library path and load required packages
```{r setup, eval =FALSE}
.libPaths("C:/Users/alaurencetraynor/Documents/2019/R/win-library/3.6")
library(aim.analysis)
library(sp)
```
#### Defaults
```{r}
name <- "Brown's Bench"
projection <- sp::CRS("+proj=aea")
```

#### Filepaths and filenames
### Filepaths
# The filepath to the folder where we have all the spatial data
```{r}
path_spatial <- "C:/Users/alaurencetraynor/Documents/2019/Analysis/BrownsBenchSiteScaleHAF/bb_sitescale_haf/Analysis/Proportional_Analysis_Request/spatial"
```
# The filepath to the folder where we have the Excel stuff

```{r}
path_tabular <- "C:/Users/alaurencetraynor/Documents/2019/Analysis/BrownsBenchSiteScaleHAF/bb_sitescale_haf/Analysis/Proportional_Analysis_Request/tabular"
```

### Filenames
```{r}
reporting_unit_filenames 

design_filenames

terradat_filename

additional_points_filename

benchmarks_filename
```


###### Read in ######
#### Reporting unit(s)
reporting_units_polygons

#### Sample designs
designs

#### TerrADat
terradat_points

#### The additional_points
additional_points

#### Benchmark Tool
benchmarks

#### Plot Status Workbook?



###### Sort out points ######
#### Prepare TerrADat points

# Also we need to get the additional points in there
# But not any points that are already in there
terradat_points

# And add reporting unit information
terradat_points

# And make sure that the fates are correct


### Add benchmark values
# First is benchmark groups
benchmark_group_lut

# Then apply the benchmarks by benchmark group
terradat_benchmarked

#### Create combined points shapefile
### Combine all sample design points
design_points

### Add plot status (if necessary)


### Replace all sampled points with their TerrADat counterparts
all_points

# There are problems with the design points relating to the additional points

# But seriously, add reporting units
all_points

###### Sort out polygons ######
#### Restrict input polygons to reporting unit(s)
strata_clipped

#### Create weight category polygons
wgtcat_polygons

#### Add weight categories to the points
all_points


###### Analyze ######
#### Generate the weights
weight_info

# Make sure the benchmark information is associated with all_points
benchmarked_points

# Analyze
analysis

###### Write results ######

```{r}

```

